---
title: "Cambridge Beer Festival 2025 Sales Figures - British Beer"
author: "Tim Rayner"
format:
  html:
    fig-width: 10
    fig-height: 8
    embed-resources: true
    anchor-sections: true
    code-fold: true
    code-overflow: scroll
    toc: true
    toc-location: left
    toc-depth: 6
    toc-expand: true
    highlight-style: arrow
    theme: simplex
    execute:
      warning: false
      cache: false
date: '2025-10-08'
---

## Introduction

This report provides a brief summary of British beer sales and costs for the 2025 Cambridge Beer Festival.

```{r setup}
library(BeerFestDB)
library(tidyverse)
library(DT)

# Edit these:
baseuri <- "https://beerfestdb.cambridgebeerfestival.uk/"
festname <- "Cambridge Beer Festival 2025"
prodcat <- "beer"
# End of editable section (but see also the credentials chunk, below).
```

```{r taglist_hack, include=FALSE}
# This is part of a hack to enable datatable in quarto tabsets; see https://stackoverflow.com/a/74708020
htmltools::tagList(DT::datatable(mtcars))
```

```{r credentials, echo=FALSE}
# To render this document using quarto, you will need to include your login details below (read-only account is fine).
# ALWAYS keep the echo=FALSE tag above, to avoid including account details in the output report.
creds <- list(username = "REDACTED", password = "REDACTED")
```

## Sales Figures

This code retrieves the dip data from the database for the festival and product category of interest:

```{r retrieve_data}
# The 'creds' list value is set in a hidden chunk to avoid exposing user account credentials.
cp <- getFestivalData(baseuri, festname, prodcat, auth = creds) %>%
  mutate(
    cask_price = as.numeric(cask_price) / 100,
    sale_price = as.numeric(sale_price) / 100
  )

# Set NA stillage to 'Unassigned'
cp$stillage[is.na(cp$stillage)] <- "Unassigned"

# Limit stillage string lenghth to 15
w <- nchar(cp$stillage) > 15
cp[w, "stillage"] <- paste(substr(cp[w, "stillage"], 0, 12), "...", sep = "")

# This assumes the dip columns are returned in order, so the rightmost column is the final dip:
w <- colnames(cp) == "cask_volume" | grepl("^dip\\.", colnames(cp))
cp$volume_lost <- cp[, w][, sum(w)]
cp$volume_lost[cp$is_condemned == 1] <- cp$cask_volume # Assume entire cask lost when condemned
cp$volume_sold <- cp$cask_volume - cp$volume_lost
write.csv(cp, "full_dip_dump.csv", row.names = FALSE)
```

Here we run the calculation of mean sales volumes and costs across container types and ABV ranges:

```{r calculation}
# Function to fess up to all-NA values but to optionally gloss over partial-NA values (unreceived invoices)
mysum <- function(x, ...) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, ...))
  }
}
agg <- cp %>%
  group_by(
    company_name, product_name, cask_volume, size_name,
    is_sale_or_return, abv, cask_price, sale_price
  ) %>%
  summarise(
    pints_sold = sum(volume_sold) * 8, # Assumes volumes in gallons
    pints_lost = sum(volume_lost) * 8,
    pints_total = sum(cask_volume) * 8,
    cost_total = mysum(cask_price, na.rm = TRUE), # na.rm=TRUE would ideally not be needed here (invoice issue)
    .groups = "drop"
  ) %>%
  mutate(
    abv_class = cut(abv, c(0, 0.5, 3, 4, 5, 6, 7.5, Inf), right = FALSE),
    sales_total = sale_price * pints_sold,
    sales_lost = sale_price * pints_lost
  ) %>%
  group_by(abv_class, is_sale_or_return, size_name) %>%
  summarise(
    pints_sold = round(sum(pints_sold), 2),
    pints_unsold = round(sum(pints_lost), 2),
    pints_total = round(sum(pints_total), 2),
    sales_total = round(sum(sales_total), 2),
    cost_total = round(mysum(cost_total, na.rm = TRUE), 2)
  ) %>%
  mutate(
    average_sale_price = ifelse(pints_sold > 0, round(sales_total / pints_sold, 2), NA),
    cost_per_pint = ifelse(pints_total > 0, round(cost_total / pints_total, 2), NA)
  ) %>%
  arrange(size_name, is_sale_or_return, abv_class) %>%
  select(-any_of(c("sales_total", "cost_total"))) %>%
  mutate(across(c(average_sale_price, cost_per_pint), function(x) sprintf("%.2f", x))) %>%
  ungroup()

levels(agg$abv_class) <- sub(",", " - ", sub("\\[", "", sub("\\)", "%", levels(agg$abv_class))))
```

Listing the results for paid casks, split by container size:

```{r paid_cask_results}
#| results: asis
x <- subset(agg, is_sale_or_return == 0) %>%
  select(-all_of(c("is_sale_or_return")))
x <- split.data.frame(x, list(x$size_name))
knitr_tabset(x, function(i) {
  show(htmltools::tagList(DT::datatable(i,
    autoHideNavigation = TRUE, rownames = FALSE,
    options = list(pageLength = 20)
  )))
})
```

And the results for sale-or-return casks:

```{r sor_cask_results}
#| results: asis
x <- subset(agg, is_sale_or_return == 1) %>%
  select(-all_of(c("is_sale_or_return")))
x <- split.data.frame(x, list(x$size_name))
knitr_tabset(x, function(i) {
  show(htmltools::tagList(DT::datatable(i,
    autoHideNavigation = TRUE, rownames = FALSE,
    options = list(pageLength = 20)
  )))
})
```
