---
title: "Cambridge Beer Festival Cellar Report"
format:
  html:
    fig-width: 10
    fig-height: 8
    embed-resources: true
    anchor-sections: true
    code-fold: true
    code-overflow: scroll
    toc: true
    toc-location: left
    toc-depth: 6
    toc-expand: true
    highlight-style: arrow
    theme: simplex
    execute:
      warning: false
      cache: false
date: last-modified
params:
  baseuri: 'https://beerfestdb.cambridgebeerfestival.uk/'
  prodcat: 'beer'
  username: ''
  password: ''
---


# Introduction

This report provides a summary of the sales and costs of beers at the Cambridge Beer Festival, based on the data in the BeerFestDB database. It includes overall sales data, order distributions, and sales by category (ABV, region, stillage location, and style).

```{r setup, include=FALSE}
# Load libraries
library(RCurl)
library(rjson)
library(RColorBrewer)
library(gplots)
library(tidyverse)
library(DT)
library(htmltools)
library(ggplot2)
library(reshape2)
library(tibble)
library(colorspace)

library(BeerFestDB)

# If dip batch names are not already set in the database, we assume the festival
# starts on the first of these days, and that dips are daily:
days <- c("Mon", "Tues", "Weds", "Thurs", "Fri", "Sat")

# This is part of a hack to enable datatable in quarto tabsets; see https://stackoverflow.com/a/74708020
htmltools::tagList(DT::datatable(mtcars))
```

```{r credentials, include=FALSE}
# To render this document using quarto, you will need to include your
# login details as part of the command line. It is recommended to pass
# the parameters as environmental variables. For example: 
#    quarto render dip_figure_analysis.qmd -P username:$BFDB_USERNAME -P password:$BFDB_PASSWORD

creds <- with(params, list(username = username, password = password))
```

```{r data_retrieval, include=FALSE}
f <- currentFestival(baseuri = params$baseuri, auth = creds)

festival <- with(params, getFestivalData(baseuri, f$name, prodcat, auth = creds, dipnames = days))

# This assumes the dip columns are returned in order, so the rightmost column is the final dip:
dip_cols <- c("cask_volume", days)

festival$write_csv("full_dip_dump.csv")
```


# Overall Sales Figures

The following table summarises the volumes sold, unsold and condemned, further broken down by whether the cask was sold on a sale-or-return basis or not.

```{r overall_sales_data}
vol.total <- sum(festival$data$cask_volume)
pct_format <- function(x) { paste0(round(x / vol.total * 100, 1), '%') }

ob <- with(festival$data,
           aggregate(data.frame(cask_volume, volume_sold, volume_lost),
                     list(order_batch, is_condemned, is_sale_or_return), sum)) %>% 
  rename(value = 'Group.1', is_condemned = 'Group.2', is_sor = 'Group.3') %>%
  mutate(value = paste('Order:', value),
         is_condemned = factor(is_condemned, levels=c(0,1)),
         is_sor = factor(is_sor, levels=c(0,1)))

d.total <- ob %>%
  summarise(value = 'Total Ordered',
            cask_volume = sum(cask_volume))

d.sor <- ob %>% filter(is_sor == 1) %>%
  summarise(value = 'Sale or Return Ordered',
            cask_volume = sum(cask_volume))

calcSales <- function(ob, tag='Total') {
  d.sales <- ob %>% 
    group_by(is_condemned, .drop = FALSE) %>%
    summarise(value = 'dummy',
              cask_volume = sum(volume_lost), ) %>%
    mutate(value = ifelse(is_condemned == 1,
                          paste0(tag, ' Condemned'),
                          paste0(tag, ' Unsold'))) %>%
    select(-any_of(c('is_condemned')))

  d.sales <- ob %>%
    filter(is_condemned == 0) %>%
    summarise(value = paste0(tag, ' Sold'),
             cask_volume = sum(volume_sold)) %>%
    rbind(d.sales)
  return(d.sales)
}

d.sales <- calcSales(ob) %>% rbind(calcSales(ob %>% filter(is_sor == 1), tag='Sale or Return'))

d <- ob %>% group_by(value) %>%
  summarise(cask_volume = sum(cask_volume)) %>%
  arrange(desc(which.max(cask_volume) == seq_len(n()))) %>%
  rbind(d.total) %>%
  rbind(d.sor) %>%
  rbind(d.sales) %>%
  mutate(kils = cask_volume / 18,
         pct_total = pct_format(cask_volume))

d <- d %>% mutate(across(where(is.double), \(x) round(x, 2)))

#show(htmltools::tagList(
    DT::datatable(d, rownames=FALSE, options=list(pageLength=nrow(d)), height=500,
                  filter = "none", selection = "none", autoHideNavigation=TRUE,
                  colnames=c('', 'Volume (gallons)', 'Volume (kils)', 'Percent Total'),
                  caption = htmltools::tags$caption(
                      style = 'caption-side: bottom; text-align: center; font-size: 90%;',
                      'Overall Sales Data'
                      )
                  )
#    )
#)
```

Here we calculate the per diem sales levels for each cask, and aggregate those figures by beer style.

```{r overall_sales_calculation, include=FALSE}
cps <- festival$subset(is_condemned=0)

dp <- cps$grouped_per_diem_sales('style')

## Drop the last part of the festival (1/4 rounded up) since it'll
## usually be non-linear by then.
dn <- ceiling((ncol(dp)-1)/4)
drop <- rev(rev(colnames(dp))[ seq_len(dn) ])
```

This bar chart shows the amount of beer sold per day, broken down by style. The colours for each style are consistent with the other plots in this report.

```{r overall_sales_plot}
d <- reshape2::melt(dp)
d$style <- factor(d$style)
cols <- darken(colorRampPalette(brewer.pal(12, 'Set3'))(nlevels(d$style)), amount=0.1)

ggplot(d, aes(x=as.numeric(variable), y=value, fill=style)) +
  geom_area(stat="identity", colour="black", alpha=0.7) + scale_fill_manual(values=cols) +
  scale_x_continuous(breaks=1:length(days), labels=days) +
  labs(x='Festival Day', y='Volume Sold (gallons)', color='Style') +
  theme_minimal() + ggtitle("Overall sales") +
  theme(legend.position='right', legend.title=element_text(size=12), legend.text=element_text(size=10))

ggplot(d, aes(x=variable, y=value, colour=style, group=style)) +
  geom_line(size=2) + scale_colour_manual(values = cols) +
  labs(x='Festival Day', y='Volume Sold (gallons)', color='Style') +
  theme_minimal() + ggtitle("Sales by Style") +
  theme(legend.position='right', legend.title=element_text(size=12), legend.text=element_text(size=10))
```


# Budget Analysis
 
This report provides a comparison of British beer sales with costs for the Cambridge Beer Festival.

## Sales Figures

Here we run the calculation of mean sales volumes and costs across container types and ABV ranges:

```{r calculation}
# Function to fess up to all-NA values but to optionally gloss over partial-NA values (unreceived invoices)
mysum <- function(x, ...) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, ...))
  }
}
agg <- festival$data %>%
  group_by(
    company_name, product_name, cask_volume, size_name,
    is_sale_or_return, abv, cask_price, sale_price
  ) %>%
  summarise(
    pints_sold = sum(volume_sold) * 8, # Assumes volumes in gallons
    pints_lost = sum(volume_lost) * 8,
    pints_total = sum(cask_volume) * 8,
    cost_total = mysum(cask_price, na.rm = TRUE), # na.rm=TRUE would ideally not be needed here (invoice issue)
    .groups = "drop"
  ) %>%
  mutate(
    abv_class = cut(abv, c(0, 0.5, 3, 4, 5, 6, 7.5, Inf), right = FALSE),
    sales_total = sale_price * pints_sold,
    sales_lost = sale_price * pints_lost
  ) %>%
  group_by(abv_class, is_sale_or_return, size_name) %>%
  summarise(
    pints_sold = round(sum(pints_sold), 2),
    pints_unsold = round(sum(pints_lost), 2),
    pints_total = round(sum(pints_total), 2),
    sales_total = round(sum(sales_total), 2),
    cost_total = round(mysum(cost_total, na.rm = TRUE), 2),
    .groups = "drop"
  ) %>%
  mutate(
    average_sale_price = ifelse(pints_sold > 0, round(sales_total / pints_sold, 2), NA),
    cost_per_pint = ifelse(pints_total > 0, round(cost_total / pints_total, 2), NA)
  ) %>%
  arrange(size_name, is_sale_or_return, abv_class) %>%
  select(-any_of(c("sales_total", "cost_total"))) %>%
  mutate(across(c(average_sale_price, cost_per_pint), function(x) sprintf("%.2f", x))) %>%
  ungroup()

levels(agg$abv_class) <- sub(",", " - ", sub("\\[", "", sub("\\)", "%", levels(agg$abv_class))))
```

Listing the results for paid casks, split by container size:

```{r paid_cask_results}
#| results: asis
#| column: page
x <- subset(agg, is_sale_or_return == 0) %>%
  select(-all_of(c("is_sale_or_return")))
x <- split.data.frame(x, list(x$size_name))
knitr_tabset(x, function(i) {
  show(htmltools::tagList(DT::datatable(i,
    autoHideNavigation = TRUE, rownames = FALSE,
    options = list(pageLength = 20)
  )))
})
```

And the results for sale-or-return casks:

```{r sor_cask_results}
#| results: asis
#| column: page
x <- subset(agg, is_sale_or_return == 1) %>%
  select(-all_of(c("is_sale_or_return")))
x <- split.data.frame(x, list(x$size_name))
knitr_tabset(x, function(i) {
  show(htmltools::tagList(DT::datatable(i,
    autoHideNavigation = TRUE, rownames = FALSE,
    options = list(pageLength = 20)
  )))
})
```


# Sales rates over time

FIXME in the following, I think we need to replace the predicted/observed bars in the plots with a simple ratio of the two, with some notes indicating that this should be as close to 1 as possible.

## By ABV

```{r sales_by_abv}
plotSalesRate( cps, 'abv_class', cols=cols )
plotModelRatio( cps, 'abv_class', drop=drop )
```

## By Region
```{r sales_by_region}
plotSalesRate( cps, 'region', cols=cols )
plotModelRatio( cps, 'region', drop=drop )
```

## By Stillage Location
```{r sales_by_stillage}
plotSalesRate( cps, 'stillage', cols=cols )
plotModelRatio( cps, 'stillage', drop=drop )
```

## By Style

The heatmap below shows the days on which each style sold best, with the colours indicating the relative sales rate for that style on that day. The bar chart below shows the ratio of predicted to observed sales for each style, with a value of 1 indicating that the beer order is predicting the sales accurately.

```{r sales_by_style}
plotSalesRate( cps, 'style', cols=cols )

d <- cps$grouped_per_diem_sales('style') %>% column_to_rownames('style')

heatmap.2(as.matrix(t(scale(t(d)))),
          dendrogram='none', key=FALSE, margins=c(5,12), trace='none',
          Colv=FALSE, lhei=c(2, 10), lwid=c(2, 5), cexRow=1.5, cexCol=1.5)

plotModelRatio( cps, 'style', drop=drop )
```


# Summary of ordered beer proportions

Here we simply show the proportion of beers ordered by geographic region and style, to give an idea of the diversity of beers ordered for the festival. FIXME perhaps don't use pie charts here? Also - why not illustrate what we ordered vs. what we sold, or vs. sales rate proportion? Why not all three? And don't scale the stacked barcharts so we get to see how much more we could have sold?

```{r order_regional_distributions}
byprod <- with(festival$data, aggregate(data.frame(region, style), list(company_name, product_name), unique))
drawPie( byprod, 'region', cols=cols, cex=1.25 ) # FIXME different colour scheme for region
drawPie( byprod, 'style', cols=cols, cex=1.25 )
```


# Technical Information

The version number of R and packages loaded for generating this
report were as follows:

```{r sessioninfo}
sessionInfo()
```
